PAGAI

### 2012-05-13 papers/TAPAS2012/pagai.tex
tapas

\usepackage{amsmath}
\usepackage{mathcomp,amsfonts,amssymb,MnSymbol}
\usepackage{natbib}
\usepackage[mathscr]{euscript}
\usepackage{algorithmic,algorithm}
%\usepackage{graphicx}
%\usepackage{color}

@@

    \thanks[imag.fr]{Email:
    \href{First Name.Last Name@imag.fr} {\texttt{\normalshape
        First Name.Last Name@imag.fr}}} 
    \thanks[grenoble-inp.fr]{Email:
    \href{Matthieu.Moy@grenoble-inp.fr} {\texttt{\normalshape
        Matthieu.Moy@grenoble-inp.fr}}} 
\begin{abstract} 
	We present design and implementation of PAGAI, a new static analyzer 
	working over the LLVM compiler infrastructure, that computes inductive
	invariants on the numerical variables of the program.
	PAGAI implements various state-of-the-art techniques combining Abstract
	Interpretation and Decision Procedures (SMT-solving), and offers
	several abstract domains.
	We thus can compare in time and precision the various combinations of
	techniques and abstract domains. We conducted extensive experiments for
	comparing these techniques on several benchmarks and GNU programs.
\end{abstract}
\begin{keyword}
  Please list keywords from your paper here, separated by commas.
\end{keyword}
\end{frontmatter}
\section{Introduction}\label{intro}
Introduction

@@

%\MM{On peut peut-être citer le nom de l'outil (on garde PAGAI ?) ici.
%  L'intérêt d'avoir un nom, c'est que les gens peuvent en parler plus
%  facilement, et c'est plus facile de retrouver le papier origine d'un
%  outil quand il cite son nom}
%\DM{Dans la version finale, oui, mais pas dans une soumission anonymisée... suffit de Googlifier pagai sur imag.fr}
%
Our tool operates over LLVM bitcode \citep{LLVM_langref,Lattner:2004:LCF:977395.977673}, which is a target for several compilers, most notably Clang (supporting C and C++) and llvm-gcc (supporting C, C++, Fortran and Ada).
Abstract domains are provided by the APRON library \citep{DBLP:conf/cav/JeannetM09}, and include convex polyhedra (from the builtin Polka ``PK'' library), octagons, intervals, and linear congruences.
For SMT-solving, our analyzer uses Yices
\cite{DBLP:conf/cav/DutertreM06} or Microsoft Z3\cite{DBLP:conf/tacas/MouraB08}.

@@

\begin{figure}[h]
  \begin{center}
    %\input{./techniques.tex}
  \end{center} 
  \vspace{-20pt}
  \caption{Comparison of the abstract values obtained by classical abstract
  interpretation (S), \emph{Guided Static
  Analysis} (G), \emph{Path-focused} technique (PF), our combined technique
  (G+PF), and its version with disjunctive invariants (DIS).
  The $\subsetneq$ bars gives the percentage of invariants stronger (more precise; smaller with respect to inclusion) with the left-side technique,
$\supsetneq$ the percentage of invariants stronger with the right-side technique,
and ``uncomparable'' gives the percentage of invariants that are uncomparable, i.e
neither greater nor smaller;
the code points where both invariants are equal make up the remaining percentage. See Tab.~\ref{tab:techniques} for details.}
  \label{fig:techniques}
\end {figure}

\begin{table*}
\begin{center}
\setlength{\tabcolsep}{0.75ex}
\begin{tabular}{|l
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}|} \hline
\multicolumn{1}{|c|}{\textbf{Benchmark}}
& \multicolumn{3}{c|}{\textbf{G/S}}
& \multicolumn{3}{c|}{\textbf{PF/S}}
& \multicolumn{3}{c|}{\textbf{PF/G}}
& \multicolumn{3}{c|}{\textbf{G+PF/PF}}
& \multicolumn{3}{c|}{\textbf{G+PF/G}}
& \multicolumn{3}{c|}{\textbf{DIS/G+PF}} \\ %\cline{2-16}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.} \\
 \hline
 %\input{table1}
\end{tabular}
\end{center}
\caption{Result of the comparison of the various techniques described in this
paper: classic Abstract Interpretation (S), \emph{Guided Static Analysis} (G),
\emph{Path-focusing} (PF), our combined technique (G+PF), and its version using
disjunctive invariants (DIS). For
instance, \textbf{G/S} compares the benefits of \emph{Guided Static Analysis}
over the classic Abstract interpretation algorithm.
$\subsetneq$, $\supsetneq$ and ``unc.'' are defined as in Fig.~\ref{fig:techniques}.}
\label{tab:techniques}
\end{table*}

@@

\begin{table}[!h]
	\centering
\begin{tabular}{|l|r|r|r|r|r|} \hline
	\multicolumn{1}{|c|}{Benchmark} &
        \multicolumn{1}{c|}{\textbf{S}} &
        \multicolumn{1}{c|}{\textbf{G}} &
        \multicolumn{1}{c|}{\textbf{PF}} &
        \multicolumn{1}{c|}{\textbf{G+PF}} &
	\multicolumn{1}{c|}{\textbf{DIS}} \\ \hline
	%\input{table_time.tex} \hline
\end{tabular}
\caption{Execution time for each technique, expressed in seconds}
\label{tab:time}
\end{table}

\subsection{Precision of Abstract Domains}
\label{sec:compare_domains}
\begin{table*}
\begin{center}
\setlength{\tabcolsep}{0.75ex}
\begin{tabular}{|l
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}|} \hline
\multicolumn{1}{|c|}{\textbf{Benchmark}}
& \multicolumn{3}{c|}{\textbf{PK/OCT}}
& \multicolumn{3}{c|}{\textbf{PK/BOX}}
& \multicolumn{3}{c|}{\textbf{OCT/BOX}}
& \multicolumn{3}{c|}{\textbf{PK/PKEQ}}
& \multicolumn{3}{c|}{\textbf{PK/GRID}}
& \multicolumn{3}{c|}{\textbf{POLY*/POLY}} \\ %\cline{2-16}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.} \\
 \hline
 %\input{table_domain.tex}
\end{tabular}
\end{center}
\caption{Results of the comparison of the various abstract domains, when using
the same technique (G+PF). We used as abstract domains Convex Polyhedra (PK and
POLY),
Octagons (OCT), intervals (BOX), linear equalities (PKEQ) and linear congruences
(GRID). The last column compares the domain of Convex Polyhedra with the
improved widening operator from \citet{BagnaraHRZ05SCP} (POLY*), and Convex Polyhedra
using the classical widening operator (POLY). POLY and POLY* use the Parma
Polyhedral Library \cite{PPL}.
$\subsetneq$, $\supsetneq$ and ``unc.'' are defined as in Fig.~\ref{fig:techniques}.}
\label{tab:domain}
\end{table*}

>>>>

\usepackage{amsmath}
\usepackage{mathcomp,amsfonts,amssymb,MnSymbol}
\usepackage{natbib}
\usepackage[mathscr]{euscript}
\usepackage{algorithmic,algorithm}
\usepackage{graphicx}
\usepackage{color}

@@

    \thanks[imag.fr]{Email:
    \href{First Name.Last Name@imag.fr} {\texttt{\normalshape
        First Name.Last Name@imag.fr}}} 
    \thanks[grenoble-inp.fr]{Email:
    \href{Matthieu.Moy@grenoble-inp.fr} {\texttt{\normalshape
        Matthieu.Moy@grenoble-inp.fr}}} 
\begin{abstract} 
	We present design and implementation of PAGAI, a new static analyzer 
	working over the LLVM compiler infrastructure, that computes inductive
	invariants on the numerical variables of the program.
	PAGAI implements various state-of-the-art techniques combining Abstract
	Interpretation and Decision Procedures (SMT-solving), and offers
	several abstract domains.
	We thus can compare in time and precision the various combinations of
	techniques and abstract domains. We conducted extensive experiments for
	comparing these techniques on several benchmarks and GNU programs.
\end{abstract}
\begin{keyword}
  Please list keywords from your paper here, separated by commas.
\end{keyword}
\end{frontmatter}
\section{Introduction}\label{intro}

This paper introduces PAGAI, a new tool for static analysis, that
computes numerical invariants at the different control points of a program.
This tool uses the LLVM compiler infrastructure.

Several other verification tools already use the LLVM framework, such as
Calysto, KLEE, LAV, LLBMC.

Our tool is dedicated to experiment various combination of analysis techniques
by abstract interpretation.
Indeed, depending on the iteration strategy, the choice of the abstract domain,
etc, precision and cost of the analysis hugely vary. Our platform allows to
compare various state-of-the-art techniques and abstract domains on real life
programs, such as GNU projects.

@@

%\MM{On peut peut-être citer le nom de l'outil (on garde PAGAI ?) ici.
%  L'intérêt d'avoir un nom, c'est que les gens peuvent en parler plus
%  facilement, et c'est plus facile de retrouver le papier origine d'un
%  outil quand il cite son nom}
%\DM{Dans la version finale, oui, mais pas dans une soumission anonymisée... suffit de Googlifier pagai sur imag.fr}
%
Our tool operates over LLVM bitcode \citep{LLVM_langref,Lattner:2004:LCF:977395.977673}, which is a target for several compilers, most notably Clang (supporting C and C++) and llvm-gcc (supporting C, C++, Fortran and Ada).
Abstract domains are provided by the APRON library \citep{DBLP:conf/cav/JeannetM09}, and include convex polyhedra (from the builtin Polka ``PK'' library), octagons, intervals, and linear congruences.
For SMT-solving, our analyzer uses Yices
\cite{DBLP:conf/cav/DutertreM06} or Microsoft Z3\cite{DBLP:conf/tacas/MouraB08}
through their C API, and can also operate with any SMT-solver supporting the
SMT-Lib 2 standard \cite{BarST-SMTLIB} through a pipe.

@@

\begin{figure}[h]
  \begin{center}
    \input{./gnuplot/techniques.tex}
  \end{center} 
  \vspace{-20pt}
  \caption{Comparison of the abstract values obtained by classical abstract
  interpretation (S), \emph{Guided Static
  Analysis} (G), \emph{Path-focused} technique (PF), our combined technique
  (G+PF), and its version with disjunctive invariants (DIS).
  The $\subsetneq$ bars gives the percentage of invariants stronger (more precise; smaller with respect to inclusion) with the left-side technique,
$\supsetneq$ the percentage of invariants stronger with the right-side technique,
and ``uncomparable'' gives the percentage of invariants that are uncomparable, i.e
neither greater nor smaller;
the code points where both invariants are equal make up the remaining percentage. See Tab.~\ref{tab:techniques} for details.}
  \label{fig:techniques}
\end {figure}

\begin{table*}
	\tiny
\begin{center}
\setlength{\tabcolsep}{0.75ex}
\begin{tabular}{|l
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}|} \hline
\multicolumn{1}{|c|}{\textbf{Benchmark}}
& \multicolumn{3}{c|}{\textbf{G/S}}
& \multicolumn{3}{c|}{\textbf{PF/S}}
& \multicolumn{3}{c|}{\textbf{PF/G}}
& \multicolumn{3}{c|}{\textbf{G+PF/PF}}
& \multicolumn{3}{c|}{\textbf{G+PF/G}}
& \multicolumn{3}{c|}{\textbf{DIS/G+PF}} \\ %\cline{2-16}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.} \\
 \hline
 \input{table1}
\end{tabular}
\end{center}
\caption{Result of the comparison of the various techniques described in this
paper: classic Abstract Interpretation (S), \emph{Guided Static Analysis} (G),
\emph{Path-focusing} (PF), our combined technique (G+PF), and its version using
disjunctive invariants (DIS). For
instance, \textbf{G/S} compares the benefits of \emph{Guided Static Analysis}
over the classic Abstract interpretation algorithm.
$\subsetneq$, $\supsetneq$ and ``unc.'' are defined as in Fig.~\ref{fig:techniques}.}
\label{tab:techniques}
\end{table*}

@@

\begin{table}[!h]
	\centering \tiny
\begin{tabular}{|l|r|r|r|r|r|} \hline
	\multicolumn{1}{|c|}{Benchmark} &
        \multicolumn{1}{c|}{\textbf{S}} &
        \multicolumn{1}{c|}{\textbf{G}} &
        \multicolumn{1}{c|}{\textbf{PF}} &
        \multicolumn{1}{c|}{\textbf{G+PF}} &
	\multicolumn{1}{c|}{\textbf{DIS}} \\ \hline
	\input{table_time.tex} \hline
\end{tabular}
\caption{Execution time for each technique, expressed in seconds}
\label{tab:time}
\end{table}

\subsection{Precision of Abstract Domains}
\label{sec:compare_domains}
\begin{table*}
	\tiny
\begin{center}
\setlength{\tabcolsep}{0.75ex}
\begin{tabular}{|l
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}%
|D{.}{.}{2}D{.}{.}{2}D{.}{.}{2}|} \hline
\multicolumn{1}{|c|}{\textbf{Benchmark}}
& \multicolumn{3}{c|}{\textbf{PK/OCT}}
& \multicolumn{3}{c|}{\textbf{PK/BOX}}
& \multicolumn{3}{c|}{\textbf{OCT/BOX}}
& \multicolumn{3}{c|}{\textbf{PK/PKEQ}}
& \multicolumn{3}{c|}{\textbf{PK/GRID}}
& \multicolumn{3}{c|}{\textbf{POLY*/POLY}} \\ %\cline{2-16}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.}
& \multicolumn{1}{c}{$\subsetneq$} & \multicolumn{1}{c}{$\supsetneq$} & \multicolumn{1}{c|}{unc.} \\
 \hline
 \input{table_domain.tex}
\end{tabular}
\end{center}
\caption{Results of the comparison of the various abstract domains, when using
the same technique (G+PF). We used as abstract domains Convex Polyhedra (PK and
POLY),
Octagons (OCT), intervals (BOX), linear equalities (PKEQ) and linear congruences
(GRID). The last column compares the domain of Convex Polyhedra with the
improved widening operator from \citet{BagnaraHRZ05SCP} (POLY*), and Convex Polyhedra
using the classical widening operator (POLY). POLY and POLY* use the Parma
Polyhedral Library \cite{PPL}.
$\subsetneq$, $\supsetneq$ and ``unc.'' are defined as in Fig.~\ref{fig:techniques}.}
\label{tab:domain}
\end{table*}
